---
title: "Untitled"
output: html_document
date: "2024-04-30"
editor_options: 
  chunk_output_type: console
---

Author: Marie Vibeke Vestergaard
This is adapted from the code from Nathan Constantine-Cooke

This script will find IBD patients eligible for this study
- They should be diagnosed after 2015

Read in packages and file paths:
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(lubridate)
path_registers <- "/ngc/projects2/predict_r/research/data/sds_registers/2022-09"
path_project <- "/ngc/projects2/predict_r/research/projects/0042_IBD_fcal_crp_trajectories"
path_basefiles <- "/ngc/projects2/predict_r/research/data/basefiles/2023-07" 
```

## Select IBD patients:
The full IBD cohort and lab results are extracted from registries previously
```{r}
full_ibd_cohort <- read.csv("/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/complementary_files/All_IBD_patients.csv", sep=";")

#Select diagnosed after 2015:
cohort <- full_ibd_cohort %>% filter(as.Date(diag_date)>=as.Date("2015-01-01"))

#Lab tests
lab_IBD <- read.csv("/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/lab/IBD_lab_Nathan.txt.gz", sep="\t")
lab_IBD <- lab_IBD %>% filter(lbnr %in% cohort$lbnr)

f_cal <- c("NPU19717", "NPU26814")
crp <- c("NPU19748", "NPU01423")

table(lab_IBD$ANALYSISCODE)
# NPU01423 NPU19717 NPU19748 NPU26814  -> (= CRP, f-cal, CRP, f-cal)
#    24815   237726  1797955    90098 

p1 <- ggplot(lab_IBD %>% filter(ANALYSISCODE %in% crp))+
  geom_density(aes(x=as.numeric(VALUE), fill=ANALYSISCODE), alpha=0.5)+
  theme_classic()+xlim(c(0,400))

ggsave(p1, file=paste0(path_project, "/Illustrations/NPU_distributions_CRP.tiff"))

p2 <- ggplot(lab_IBD %>% filter(ANALYSISCODE %in% f_cal))+
  geom_density(aes(x=as.numeric(VALUE), fill=ANALYSISCODE), alpha=0.5)+
  theme_classic()+xlim(c(0,5000))

ggsave(p2, file=paste0(path_project, "/Illustrations/NPU_distributions_Fcal.tiff"))

#FC values appear similar for both NPU codes - merge!
#CRP values appears to be higher for NPU01423. It has much fewer samples --> for consistency, just go for just NPU19748
crp <- c("NPU19748")

#Combine and filter
#Period: from 90 days before diagnosis till 5 years after. 
lab_IBD <- left_join(lab_IBD, full_ibd_cohort %>% select(lbnr, diag_date, final_diag, sex, birth_date), by="lbnr") %>% filter(as.Date(SAMPLINGDATE)>=as.Date(diag_date)-days(90)) %>% filter(as.Date(SAMPLINGDATE)<=as.Date(diag_date)+years(5))

lab_crp <- lab_IBD %>% filter(ANALYSISCODE %in% crp)
lab_fcal<- lab_IBD %>% filter(ANALYSISCODE %in% f_cal)

#CRP
#Those with sampling at diagnosis +- 90 days
crp_diag <- lab_crp %>% filter(as.Date(SAMPLINGDATE)>=as.Date(diag_date)-days(90) & as.Date(SAMPLINGDATE)<=as.Date(diag_date)+days(90))

#Those with at least 3 in total:
crp_rest <- lab_crp %>% group_by(lbnr) %>% mutate(n=n()) %>% filter(n>2)

lab_crp <- lab_crp %>% filter(lbnr %in% crp_diag$lbnr & lbnr %in% crp_rest$lbnr)
length(unique(lab_crp$lbnr)) # 7740 IBD patients 

lab_crp <- lab_crp %>% mutate(final_diag = ifelse(final_diag %in% c("CD", "CD_unc"), "CD", "UC"))

#F-cal:
#Those with sampling at diagnosis
fcal_diag <- lab_fcal %>% filter(as.Date(SAMPLINGDATE)>=as.Date(diag_date)-days(90) & as.Date(SAMPLINGDATE)<=as.Date(diag_date)+days(90))

#Those with at least 3 in total:
fcal_rest <- lab_fcal %>% group_by(lbnr) %>% mutate(n=n()) %>% filter(n>2)

#Those fulfilling both criteria:
lab_fcal <- lab_fcal %>% filter(lbnr %in% fcal_diag$lbnr & lbnr %in% fcal_rest$lbnr)
length(unique(lab_fcal$lbnr)) #5445 IBD patients 

lab_fcal <- lab_fcal %>% mutate(final_diag = ifelse(final_diag %in% c("CD", "CD_unc"), "CD", "UC"))

#Save files:
write.table(lab_crp, paste0(path_project, "/GeneratedData/CRP_values.txt"), sep="\t", quote = F, row.names = F, col.names = T)
write.table(lab_fcal, paste0(path_project, "/GeneratedData/Fcal_values.txt"), sep="\t", quote = F, row.names = F, col.names = T)
```

## Retiming of sampling
```{r}
lab_crp <- read.csv(paste0(path_project, "/GeneratedData/CRP_values.txt"), sep="\t") %>% mutate(diag_date = as.Date(diag_date), SAMPLINGDATE = as.Date(SAMPLINGDATE), birth_date = as.Date(birth_date))
lab_fcal <- read.csv(paste0(path_project, "/GeneratedData/Fcal_values.txt"), sep="\t")%>% mutate(diag_date = as.Date(diag_date), SAMPLINGDATE = as.Date(SAMPLINGDATE), birth_date = as.Date(birth_date))

#Check, that there are no missing date of birth, sex or date of diagnosis
sum(is.na(lab_crp$birth_date))
sum(is.na(lab_crp$sex))
sum(is.na(lab_crp$diag_date))
sum(is.na(lab_fcal$birth_date))
sum(is.na(lab_fcal$sex))
sum(is.na(lab_fcal$diag_date))

# Add age at IBD diagnosis
lab_crp <- lab_crp %>% mutate(diag_age = as.numeric(diag_date-birth_date)/365.24)
lab_fcal <- lab_fcal %>% mutate(diag_age = as.numeric(diag_date-birth_date)/365.24)

#Retiming: If there's a test in the 90 days prior to diagnosis, this one's sampling date will be t=0. Otherwise, t=0 is at date of diagnosis
lab_crp <- lab_crp %>% arrange(SAMPLINGDATE) %>% group_by(lbnr) %>% mutate(first_sampling = min(SAMPLINGDATE))
lab_crp <- lab_crp %>% mutate(new_diag_date = as.Date(ifelse(first_sampling<diag_date, first_sampling, diag_date))) %>% mutate(t = as.numeric(SAMPLINGDATE-new_diag_date)) %>% arrange(lbnr, SAMPLINGDATE)

lab_fcal <- lab_fcal %>% arrange(SAMPLINGDATE) %>% group_by(lbnr) %>% mutate(first_sampling = min(SAMPLINGDATE))
lab_fcal <- lab_fcal %>% mutate(new_diag_date = as.Date(ifelse(first_sampling<diag_date, first_sampling, diag_date))) %>% mutate(t = as.numeric(SAMPLINGDATE-new_diag_date)) %>% arrange(lbnr, SAMPLINGDATE)

#Reapply filter of <= 5 years and >=3 measurements:
lab_crp <- lab_crp %>% filter(t<=5*365.24) %>% group_by(lbnr) %>% mutate(n=n()) %>% filter(n>=3)
length(unique(lab_crp$lbnr)) #14186 IBD patients

lab_fcal <- lab_fcal %>% filter(t<=5*365.24) %>% group_by(lbnr) %>% mutate(n=n()) %>% filter(n>=3)
length(unique(lab_fcal$lbnr)) #8762 IBD patients
```

## Repeating Nathan's plots (LIBDR) evaluating diagnosis date:
```{r}
# Year of diagnosis (the original)
p1 <- ggplot(lab_crp)+
  geom_histogram(aes(x=year(diag_date)), fill = "#B8F2E6", color = "black", binwidth = 1)+
  theme_minimal() +
  xlab("Year of IBD diagnosis") +
  ylab("Frequency")

ggsave(p1, file=paste0(path_project, "/Illustrations/Diag_year.tiff"))

p2 <- ggplot(lab_crp, aes(x = as.factor(month(diag_date, label = TRUE)))) +
  geom_bar(color = "black", fill = "#FEC601", linewidth = 0.3) +
  theme_minimal() +
  ylab("Frequency") +
  xlab("Month of IBD Diagnosis")

ggsave(p2, file=paste0(path_project, "/Illustrations/Diag_month.tiff"))
```


## Dealing with right and left cencored test values - f-cal:
```{r}
#F-cal
lab_fcal$value <- as.numeric(lab_fcal$VALUE)
lab_fcal$VALUE[which(is.na(lab_fcal$value))] #We should recode >1800 to 1800 
lab_fcal$value <- ifelse(lab_fcal$VALUE %in% c(">1800", ">3600"), 1800, lab_fcal$value)

p1 <- ggplot(lab_fcal, aes(x = value)) +
  geom_density(linewidth = 0.8, alpha = 0.5, fill = "#9FD8CB", color = "#517664") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  xlab("FCAL (µg/g)") +
  ylab("Density")

ggsave(p1, file=paste0(path_project, "/Illustrations/Fcal_values.tiff"))

#Change all >1800 results to 1800
lab_fcal$value <- ifelse(lab_fcal$value > 1800, 1800, lab_fcal$value)

p2 <- ggplot(lab_fcal, aes(x = value)) +
  geom_density(linewidth = 0.8, alpha = 0.5, fill = "#9FD8CB", color = "#517664") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  xlab("FCAL (µg/g)") +
  ylab("Density")

ggsave(p2, file=paste0(path_project, "/Illustrations/Fcal_values_right_censor.tiff"))

#Log transformation:
p3 <- ggplot(lab_fcal, aes(x = log(value+1))) +
  geom_density(linewidth = 0.8, alpha = 0.5, fill = "#9FD8CB", color = "#517664") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  xlab("log(FCAL (µg/g))") +
  ylab("Density")

ggsave(p3, file=paste0(path_project, "/Illustrations/Fcal_values_log.tiff"))

#Plot number of f-cal measurements pr. individual
p4 <- ggplot(lab_fcal, aes(x = n)) +
  geom_histogram(binwidth = 1, fill = "#887C9B") +
  geom_vline(xintercept = 2.5, color = "#D80032") +
  theme_minimal() +
  xlab("Number of FCAL observations per subject") +
  ylab("Count")

ggsave(p4, file=paste0(path_project, "/Illustrations/Fcal_measurements_pr_individuals.tiff"))

#Timing of sampling
p5 <- ggplot(lab_fcal, aes(x = t, y = log(value+1), color = factor(lbnr))) +
  geom_line(alpha = 0.2) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  scale_color_manual(values = viridis::viridis(length(unique(lab_fcal$lbnr)))) +
  guides(color = "none") +
  xlab("Time (years)") +
  ylab("Log(FCAL (µg/g))") +
  ggtitle("")

ggsave(p5, file=paste0(path_project, "/Illustrations/Fcal_measurements_timing.tiff")) 

#Follow-up:
p6 <- ggplot(lab_fcal %>% arrange(desc(t)) %>% group_by(lbnr) %>% slice(1), aes(x = t/365)) +
  geom_histogram(color = "black", fill = "#5BBA6F", binwidth = 1) +
  theme_minimal() +
  ylab("Count") +
  xlab("Followup (years)")

ggsave(p6, file=paste0(path_project, "/Illustrations/Fcal_followup.tiff"))

#Followup vs. N
p7 <- ggplot(lab_fcal %>% arrange(desc(t)) %>% group_by(lbnr) %>% slice(1), aes(y = n, x = t/365)) +
  geom_point(color = "#FF4F79", size = 2) +
  ylab("Number of FCAL measurements") + xlab("Followup")+
  theme_minimal()

ggsave(p7, file=paste0(path_project, "/Illustrations/Fcal_followup_N.tiff"))

#Proportion of right censoring
p8 <- ggplot(lab_fcal %>% group_by(lbnr) %>% summarise(prop = sum(value == 1800)/n(), n=sum(value == 1800)),aes(
    x = prop, y = n)) + 
  geom_point(alpha = 0.5, size = 0.9, color = "#D5896F") +
  ylab("Total number of observations") +
  xlab("Proportion of right-censored observations") +
  theme_minimal() +
  theme(legend.position = "none")

ggsave(p8, file=paste0(path_project, "/Illustrations/Fcal_right_censor.tiff"))

# Exclude patients with less that three non-censored observations:
lab_fcal <- lab_fcal %>% group_by(lbnr) %>% mutate(n_cens=sum(value == 1800)) %>% filter(n-n_cens >=3)
length(unique(lab_fcal$lbnr)) #7880 IBD patients
```

## Dealing with right and left cencored test values - CRP:
```{r}
#F-cal
lab_crp$value <- as.numeric(lab_crp$VALUE)
unique(lab_crp$VALUE[which(is.na(lab_crp$value))]) # <4 --> recode to 4
lab_crp$value <- ifelse(lab_crp$VALUE == "<4" , 4, lab_crp$value)

p1 <- ggplot(lab_crp, aes(x = value)) +
  geom_density(linewidth = 0.8, alpha = 0.5, fill = "#FFC800", color = "#FF8427") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  xlab("CRP (mg/L)") +
  ylab("Density")

ggsave(p1, file=paste0(path_project, "/Illustrations/CRP_values.tiff"))


#Log transformation:
p3 <- ggplot(lab_crp, aes(x = log(value+1))) +
  geom_density(linewidth = 0.8, alpha = 0.5, fill = "#FFC800", color = "#FF8427") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  xlab("log(CRP (mg/L))") +
  ylab("Density")

ggsave(p3, file=paste0(path_project, "/Illustrations/CRP_values_log.tiff"))

#Plot number of measurements pr. individual
p4 <- ggplot(lab_crp, aes(x = n)) +
  geom_histogram(binwidth = 1, fill = "#0081A7") +
  geom_vline(xintercept = 2.5, color = "#FF88DC") +
  theme_minimal() +
  xlab("Number of CRP observations per subject") +
  ylab("Count")

ggsave(p4, file=paste0(path_project, "/Illustrations/CRP_measurements_pr_individuals.tiff"))
#Again, this looks different, as we have already applied the filtering

#Timing of sampling
p5 <- ggplot(lab_crp, aes(x = t, y = log(value+1), color = factor(lbnr))) +
  geom_line(alpha = 0.2) +
  geom_point(alpha = 0.3) +
  theme_minimal() +
  scale_color_manual(values = viridis::viridis(length(unique(lab_crp$lbnr)))) +
  guides(color = "none") +
  xlab("Time (years)") +
  ylab("Log(CRP (mg/L))") +
  ggtitle("")

ggsave(p5, file=paste0(path_project, "/Illustrations/CRP_measurements_timing.tiff"))

#Follow-up:
p6 <- ggplot(lab_crp %>% arrange(desc(t)) %>% group_by(lbnr) %>% slice(1), aes(x = t/365)) +
  geom_histogram(color = "black", fill = "#B38D97", binwidth = 1) +
  theme_minimal() +
  ylab("Count") +
  xlab("Followup (years)")

ggsave(p6, file=paste0(path_project, "/Illustrations/CRP_followup.tiff"))

#Followup vs. N
p7 <- ggplot(lab_crp %>% arrange(desc(t)) %>% group_by(lbnr) %>% slice(1), aes(y = n, x = t/365)) +
  geom_point(color = "#216869", size = 2) +
  ylab("Number of CRP measurements") + xlab("Followup")+
  theme_minimal()

ggsave(p7, file=paste0(path_project, "/Illustrations/CRP_followup_N.tiff"))

#There are many CRP measurements! Group into time intervals and use median for several observations:
id.list <- unique(lab_crp$lbnr) #14186
crp.ma <- matrix(NA, nrow = length(id.list), ncol = 5)

for (i in seq_along(id.list)) {
  subject_data <- subset(lab_crp, lbnr == id.list[i])
  for (j in seq(0, 4)) {
    if (j == 4) {
      sub.obs <- subset(
        subject_data,
        t/365 >= j - 0.5 & t/365 <= j + 1
      )
    } else {
      sub.obs <- subset(
        subject_data,
        t/365 >= j - 0.5 & t/365 < j + 0.5
      )
    }
    if (nrow(sub.obs) > 0) {
      crp.ma[i, j + 1] <- median(sub.obs$value)
    }
  }
}

rownames(crp.ma) <- id.list

crp.ma  <- crp.ma [rowSums(is.na(crp.ma)) <= 2, ] #Should have at least 3 measurements, after splitting intp intervals.

id.list <- rownames(crp.ma) #10041

crp.ma <- reshape2::melt(t(crp.ma ), id.vars = row.names(crp.ma ), na.rm = TRUE)
head(crp.ma)
colnames(crp.ma) <- c("t", "lbnr", "value")
crp.ma  <- crp.ma[, c(2, 3, 1)]
crp.ma$t <- crp.ma$t - 1

# Take into account uneven spacing at start and end
crp.ma$t <- plyr::mapvalues(crp.ma$t,
  from = c(0, 4),
  to = c(0.25, 4.25)
)

# Add diagnosis type back
length(unique(crp.ma$lbnr)) #10041 IBD patients, 21910 time points
lab_crp <- left_join(crp.ma, lab_crp %>% group_by(lbnr) %>% slice(1) %>% select(lbnr, diag_date, final_diag, sex, birth_date, diag_age, new_diag_date, n), by="lbnr")

```

Save data:
```{r}
write.table(lab_crp, paste0(path_project, "/GeneratedData/CRP_values_prep.txt"), sep="\t", quote = F, row.names = F, col.names = T)
write.table(lab_fcal %>% select(lbnr, value, t, diag_date, final_diag, sex, birth_date, diag_age, new_diag_date, n), paste0(path_project, "/GeneratedData/Fcal_values_prep.txt"), sep="\t", quote = F, row.names = F, col.names = T)
```

# IBD subtype information
```{r}
lab_one <- lab_crp %>% group_by(lbnr) %>% slice(1)
table(lab_one$final_diag)
#  CD   UC 
#4415 5626

lab_one <- lab_fcal %>% group_by(lbnr) %>% slice(1)
table(lab_one$final_diag)
#  CD   UC 
#3931 3949 
```

# Overview of f-cal and CRP over the years:
```{r}
lab_IBD <- read.csv("/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/lab/IBD_lab_Nathan.txt.gz", sep="\t")
f_cal <- c("NPU19717", "NPU26814")
crp <- c("NPU19748")

lab_crp <- lab_IBD %>% filter(ANALYSISCODE %in% crp)
lab_fcal <- lab_IBD %>% filter(ANALYSISCODE %in% f_cal)
rm(lab_IBD)

p1 <- ggplot(lab_crp) +
  geom_histogram(aes(x=year(SAMPLINGDATE)), binwidth = 1, boundary = 0.5, fill="lightblue", color="black")+
  theme_classic()+
  xlab("Year of sampling")+ylab("Number of CRP tests for patients with IBD")+
  theme(text = element_text(family="sans", size=8), axis.title = element_text(family="sans", size=8), axis.text = element_text(family="sans", size=8))+
  geom_vline(aes(xintercept=2014.5), linetype="dashed", color="red")


p2 <- ggplot(lab_fcal) +
  geom_histogram(aes(x=year(SAMPLINGDATE)), binwidth = 1, boundary = 0.5, fill="gold", color="black")+
  theme_classic()+
  xlab("Year of sampling")+ylab("Number of f-calpro tests for patients with IBD")+
  theme(text = element_text(family="sans", size=8), axis.title = element_text(family="sans", size=8), axis.text = element_text(family="sans", size=8))+
  geom_vline(aes(xintercept=2014.5), linetype="dashed", color="red")

ggsave(p1, file=paste0(path_project, "/Illustrations/CRP_total_measurements.tiff"))
ggsave(p2, file=paste0(path_project, "/Illustrations/F-calpro_total_measurements.tiff"))
```

