---
title: "Untitled"
output: html_document
date: "2024-04-30"
editor_options: 
  chunk_output_type: inline
---

Author: Marie Vibeke Vestergaard
This is adapted from the code from Nathan Constantine-Cooke

Read in packages and paths:
```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(lcmm)
library(splines)
library(kml)
library(mice)
library(patchwork)
library(ggdist)
library(ggalluvial)
library(pander)
library(qqplotr)

path_registers <- "/ngc/projects2/predict_r/research/data/sds_registers/2022-09"
path_project <- "/ngc/projects2/predict_r/research/projects/0042_IBD_fcal_crp_trajectories"
path_basefiles <- "/ngc/projects2/predict_r/research/data/basefiles/2023-07" 
```

## Find optimal number of knots in a 4-cluster model:
Search from 2-4
Run the 2_crp_m1.R script
Run the 2_crp_lcmm.R script

Read in data:
```{r}
crp <- read.csv(paste0(path_project, "/GeneratedData/CRP_values_prep.txt"), sep="\t")
crp$value <- log(crp$value+1)
crp$crp_time <- crp$t

```

## 2 knots:
```{r}
#| label: fig-2-knots
#| fig-cap: "Cluster profiles for the four-cluster model assuming two knots. The vertical teal lines indicate knot location. Mean cluster profiles are denoted by the red curves"

knots.2 <- readRDS(paste0(path_project, "/GeneratedData/cache/crp-2knots-4.RDS"))
knots.2.list <- list()
knots.2.list[[4]] <- knots.2

cairo_pdf(paste0(path_project, "/Illustrations/2-knots-CRP_4.pdf"), width = 9, height = 9)
spaghettiPlot(crp, knots.2.list, G = 4, log = TRUE, tmax = 5, sizes = T, knots = TRUE, n.knots = 2, var.time = "crp_time", knot.type = "equal")
invisible(dev.off())

spaghettiPlot(crp, knots.2.list, G = 4, log = TRUE, tmax = 5, sizes = T, knots = TRUE, n.knots = 2, var.time = "crp_time", knot.type = "equal")


```

## 3 knots
```{r}
#| label: fig-3-knots
#| fig-cap: "Cluster profiles for the four-cluster model assuming three knots. The vertical teal lines indicate knot location. Mean cluster profiles are denoted by the red curves"

knots.3 <- readRDS(paste0(path_project, "/GeneratedData/cache/crp-3knots-4.RDS"))
knots.3.list <- list()
knots.3.list[[4]] <- knots.3

cairo_pdf(paste0(path_project, "/Illustrations/3-knots-CRP_4.pdf"), width = 9, height = 9)
spaghettiPlot(crp, knots.3.list, G = 4, log = TRUE, tmax = 5, sizes = TRUE, knots = TRUE, n.knots = 3, var.time = "crp_time", knot.type = "equal")
invisible(dev.off())

spaghettiPlot(crp, knots.3.list, G = 4, log = TRUE, tmax = 5, sizes = TRUE, knots = TRUE, n.knots = 3, var.time = "crp_time", knot.type = "equal")

```


## 4 knots
```{r}
#| label: fig-4-knots
#| fig-cap: "Cluster profiles for the four-cluster model assuming four knots. The vertical teal lines indicate knot location. Mean cluster profiles are denoted by the red curves"

knots.4 <- readRDS(paste0(path_project, "/GeneratedData/cache/crp-4knots-4.RDS"))
knots.4.list <- list()
knots.4.list[[4]] <- knots.4

cairo_pdf(paste0(path_project, "/Illustrations/4-knots-CRP_4.pdf"), width = 9, height = 9)
spaghettiPlot(crp, knots.4.list, G = 4, log = TRUE, tmax = 5, sizes = T, knots = TRUE, n.knots = 4, var.time = "crp_time", knot.type = "equal")
invisible(dev.off())

spaghettiPlot(crp, knots.4.list, G = 4, log = TRUE, tmax = 5, sizes = T, knots = TRUE, n.knots = 4, var.time = "crp_time", knot.type = "equal")
```


## Model statistics
```{r}
#| results: "hold"
#| label: tbl-knot-stats
#| tbl-cap: Model statistics for differing numbers of knots for natural cubic splines


knots.df <- tibble(Knots = c(2, 3, 4),
                   loglik = c(knots.2$loglik,
                              knots.3$loglik,
                              knots.4$loglik),
                   AIC = c(knots.2$AIC,
                           knots.3$AIC,
                           knots.4$AIC),
                   BIC = c(knots.2$BIC,
                           knots.3$BIC,
                           knots.4$BIC))
knitr::kable(knots.df,
             col.names = c("Knots",
                           "Maxmum log likelihood",
                           "AIC",
                           "BIC"))
```

We have elected to use natural cubic splines with 2 knots placed at quantiles. 



## Based on 2 knots, validate the 8 clusters:
Run the 3_crp_lcmm.R script, that is using the m1 file with two knots:
time

## Load the model

```{r}
# set the number of groups
models.crp <- list()
file.name <- paste0(path_project, "/GeneratedData/cache/crp-2knots-8.RDS")
models.crp[[8]] <- readRDS(file.name)
```


## Prosterior classification:
```{r}
#| results: "asis"
#postprob_quiet <- quietly(postprob)
  cat(paste0("##### G = ", 8, "\n"))
  output <- postprob(models.crp[[8]])$result
  cat(pander::pander(output[[1]]))
  cat(pander::pander(output[[2]]))
  cat(pander::pander(output[[3]]))

```

## Residual plots:
```{r}
#| results: "asis"

  cat(paste0("##### G = ", 8, "\n"))
  plot(models.crp[[8]], shades = TRUE)

```

## Chosen model:
```{r}
#| label: fig-NCS-norm
#| fig-cap: "Plots assessing normality for the distribution of residuals for the chosen NCS model for FCAL. (A) Histrogram; (B) Q-Q plot."

p1 <- data.frame(residuals = resid(models.crp[[8]])) %>%
  ggplot(aes(x = residuals)) +
  geom_histogram(aes(y = after_stat(density)),
                 fill = "#D8829D",
                 color = "#AF6A80",
                 bins = 30) +
  geom_density(color = "#023777", size = 1.2) +
  theme_classic() +
  theme(axis.line = element_line(colour = "gray")) +
  ylab("Density") +
  xlab("Residuals") +
  ggtitle("A")

p2 <- data.frame(residuals = resid(models.crp[[8]])) %>%
  ggplot(aes(sample = residuals)) +
    stat_qq_band() +
    stat_qq_line(color = "#D8829D") +
    stat_qq_point(color = "#023777")  +
    theme_classic() +
    theme(axis.line = element_line(colour = "gray")) + 
    ylab("Theoretical Quantiles") +
    xlab("Sample Quantiles") +
  ggtitle("B")

ggsave(paste0(path_project, "/Illustrations/Residual-plot-crp-ncs.pdf"),
       plot = p1 / p2,
       width = 8,
       height = 8,
       units = "in")
print(p1 / p2)
```

```{r}
#| label: fig-NCS-norm-cluster
#| fig-cap: "Q-Q plots assessing normality for the distribution of residuals stratifed by cluster assignment."

predClass <- models.crp[[8]]$pprob[, c("lbnr", "class")]
temp <- merge(models.crp[[8]]$pred, predClass, by = "lbnr")
labels <- c("A", "B", "C", "D", "E", "F", "G", "H")
p <- list()

for (i in 1:8) {
  p[[i]] <- subset(temp, class == i) %>%
    ggplot(aes(sample = resid_ss)) +
    stat_qq_band() +
    stat_qq_line(color = "#D8829D") +
    stat_qq_point(color = "#023777")  +
    theme_classic() +
    theme(axis.line = element_line(colour = "gray")) + 
    ylab("Theoretical Quantiles") +
    xlab("Sample Quantiles") +
  ggtitle(labels[i])
}

ArrangedPlot <-  (p[[1]] + p[[2]]) / (p[[3]] + p[[4]]) / (p[[5]] + p[[6]])/ (p[[7]] + p[[8]])

ggsave(paste0(path_project,"/Illustrations/cluster-resids-crp-ncs.pdf"),
       ArrangedPlot,
       width = 8,
       height = 8,
       units = "in")
print(ArrangedPlot)
```

